Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-11 docs/proposals/NSF/ProjDesc/ProjDesc.tex
NM

 Figure \ref{fig1} shows preliminary results of multiple bodies undergoing mechanical erosion in Stokes flow. The figure shows 80 bodies of initially random size and position, undergoing erosion. This figure extends previously reported 50-body simulations to 80 bodies \cite{Quaife2018} due to a recently developed barycentric interpolation scheme \tocite, which enables closer contact between bodies and thus more bodies.
The color shows the (normalized) vorticity field surrounding the bodies. Vorticity, since it reduces to shear on solid interfaces, provides a convenient way to visualize local erosion rates, as well as the local flow intensity. Observe that erosion not only reduces the size of solid bodies, but also alters their shapes substantially. The bodies tend to become somewhat polygonal: corners develop connected by relatively flat faces. The number of faces does not appear to be easily predicted, but rather depends on the interaction with neighboring bodies as mediated by the Stokes flow. Notice that, as erosion proceeds, relatively straight channels tend to develop between the bodies. Certain channels that are initially larger can transmit more flow, which promotes local erosion rates and further widens these selected channels. In this way, erosion creates a runaway process in which small differences in initial channel size become amplified. The channelization process is made more evident by visualization of the velocity magnitude, as in Fig.~\ref{fig2}.

@@

@@

The creation and reinforcement of channels significantly impacts the macroscopic properties of the porous medium, as the PI and collaborators have just begun to examine. Fig.~\ref{fig3} shows preliminary results on how media properties change in time. First, the left-most plot shows the rate at which the area fraction decreases as the bodies erode. Time is normalized by the value $t_f$ at which all bodies vanish and hence the area fraction is zero. Notice that the rate of area reduction does not change significantly during the simulation. The middle plot shows the resistivity, or inverse permeability of the medium. In the simulations, resistivity is calculated by measuring the total flux and the pressure upstream and downstream, then using Darcy's relation \eqref{DarcyEq} to infer $k$. Naturally, as bodies erode, the resistivity they provide decreases in time. Note, however, that the vertical axis is logarithmic. Hence, the initial linear decrease corresponds to an exponential rate of resistivity reduction. The much more mild decrease in area fraction is not sufficient to account for this exponential rate, and hence the reshaping process, in particular the formation of channels, plays a pivotal role.

>>>>

@@

 Figure \ref{fig1} shows preliminary results of multiple bodies undergoing mechanical erosion in Stokes flow. The figure shows 80 bodies of initially random size and position, undergoing erosion. This figure extends previously reported 50-body simulations to 80 bodies \cite{Quaife2018} due to a recently developed barycentric interpolation scheme \citep{bar2014, bar-wu-vee2015}, which enables closer contact between bodies and thus more bodies.
The color shows the (normalized) vorticity field surrounding the bodies. Vorticity, since it reduces to shear on solid interfaces, provides a convenient way to visualize local erosion rates, as well as the local flow intensity. Observe that erosion not only reduces the size of solid bodies, but also alters their shapes substantially. The bodies tend to become somewhat polygonal: corners develop connected by relatively flat faces. The number of faces does not appear to be easily predicted, but rather depends on the interaction with neighboring bodies as mediated by the Stokes flow. Notice that, as erosion proceeds, relatively straight channels tend to develop between the bodies. Certain channels that are initially larger can transmit more flow, which promotes local erosion rates and further widens these selected channels. In this way, erosion creates a runaway process in which small differences in initial channel size become amplified. The channelization process is made more evident by visualization of the velocity magnitude, as in Fig.~\ref{fig2}.

@@

 \vsp{5} HERE

The creation and reinforcement of channels significantly impacts the macroscopic properties of the porous medium. Fig.~\ref{fig3} shows preliminary results on how media properties change in time. First, the left-most plot shows the rate at which the area fraction decreases as the bodies erode. Time is normalized by the value $t_f$ at which all bodies vanish and hence the area fraction is zero. Notice that the rate of area reduction does not change significantly during the simulation. The middle plot shows the resistivity, or inverse permeability of the medium. In the simulations, resistivity is calculated by measuring the total flux and the pressure upstream and downstream, then using Darcy's relation \eqref{DarcyEq} to infer $k$. Naturally, as bodies erode, the resistivity they provide decreases in time. Note, however, that the vertical axis is logarithmic. Hence, the initial linear decrease corresponds to an exponential rate of resistivity reduction. The much more mild decrease in area fraction is not sufficient to account for this exponential rate, and hence the reshaping process, in particular the formation of channels, plays a pivotal role.
